### 2019-06-20 README.Rmd
typo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "How to use replication assignments for teaching integrity in empirical archaeology"
```

# Teaching replication in archaeology
<!-- badges: start -->
[![Launch Rstudio Binder](http://mybinder.org/badge.svg)](https://mybinder.org/v2/gh/benmarwick/replicationteachingarchaeology/master?urlpath=rstudio) 
[![Travis-CI Build Status](https://travis-ci.org/benmarwick/teaching-replication-in-archaeology.svg?branch=master)](https://travis-ci.org/benmarwickteaching-replication-in-archaeology)
[![ORCiD](https://img.shields.io/badge/ORCiD-0000--0001--7879--4531-green.svg)](http://orcid.org/0000-0001-7879-4531)

<!-- badges: end -->

This repository contains the data and code for our paper:

> Marwick, B., L. Wang, R. Robinson, H. Loiselle (2019). _`r Title`_. Name of journal/book <https://doi.org/10.17605/OSF.IO/DBSW9>

Our pre-print is online here:

> Authors, (2019). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.17605/OSF.IO/DBSW9>

### How to cite

Please cite this compendium as:

> Marwick, B., L. Wang, R. Robinson, H. Loiselle, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for "`r Title`"_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.17605/OSF.IO/DBSW9>

### How to download or install

You can download the compendium as a zip from from this URL: [master.zip](/archive/master.zip)

Or you can install this compendium as an R package, teachingreplicationinarchaeology, from GitHub with:

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.