### 2019-08-15 README.Rmd
improve the readme to make it more clear what this is, and what to do with it

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "How to use replication assignments for teaching integrity in empirical archaeology"
```

# Teaching replication in archaeology
<!-- badges: start -->
[![Launch Rstudio Binder](http://mybinder.org/badge.svg)](https://mybinder.org/v2/gh/benmarwick/replicationteachingarchaeology/master?urlpath=rstudio) 
[![Travis-CI Build Status](https://travis-ci.org/benmarwick/teaching-replication-in-archaeology.svg?branch=master)](https://travis-ci.org/benmarwick/teaching-replication-in-archaeology)

<!-- badges: end -->

This repository contains the data and code for our paper:

> Marwick, B. [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7879-4531), L. Wang [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-9060-2125), R. Robinson, H. Loiselle (2019). _`r Title`_. Submitted to _Advances in Archaeological Practice_ <https://doi.org/10.17605/OSF.IO/DBSW9>

Our pre-print is online here:

> Marwick, B., L. Wang, R. Robinson, H. Loiselle , (2019). _`r Title`_. SocArXiv, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://osf.io/preprints/socarxiv/tsxbv>

### How to cite

Please cite this compendium as:

> Marwick, B., L. Wang, R. Robinson, H. Loiselle, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for "`r Title`"_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.17605/OSF.IO/DBSW9>

### Overview of contents

## Contents

The **analysis** directory contains:

- [:file_folder: paper](/analysis/paper): R Markdown source document for manuscript. Includes code to reproduce the figures and tables generated by the analysis. It also has a rendered version, `paper.docx`, suitable for reading (the code is replaced by figures and tables in this file)    
- [:file_folder: data](/analysis/data): Data used in the analysis.  
- [:file_folder: figures](/analysis/figures): Plots and other illustrations

### How to run in your broswer or download and run locally

The simplest way to explore the text, code and data is to click on [binder](https://mybinder.org/v2/gh/benmarwick/replicationteachingarchaeology/master?urlpath=rstudio) to open an instance of RStudio in your browser, which will have the compendium files ready to work with. Binder uses rocker-project.org Docker images to ensure a consistent and reproducible computational environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL: [master.zip](/archive/master.zip). After unzipping, open the `Proj` file in RStudio, and run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the [DESCRIPTION](/DESCRIPTION) file). Then open `paper.Rmd` and knit to produce the `paper.docx`.

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.