### 2019-08-17 analysis/supplementary-materials/assignment-instructions/template-for-archy-483-replication-report.Rmd
resubmit after peer reviews

---
title: "Spring 2019 ARCHY 483 Replication Report"
author: "My Name"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # this ensures we can see your code in the output
```

## Introduction

## Methods

## Results

```{r read-in-the-data}

```

```{r tidy-the-data}

```

```{r plot-the-data}

```

## Conclusion

## References cited

<!-- don't delete the line below, it will count the words in your report. You will need to do a one-time install of the wordcountaddin package first by running this code in your console: source("https://install-github.me/benmarwick/wordcountaddin") #  -->
Word count: `r wordcountaddin::word_count()`